---
title: "Data_for_RR"
author: "MLB"
date: "2024-07-22"
output: html_document
---

# Packages

```{r}
# Install 
## Run 1x !!!

# install.packages("flowCore")
# install.packages("flowWorkspace")
# install.packages("openCyto")
# install.packages("flowAI")
# install.packages("gridExtra")
# install.packages("tidyverse)
```

```{r}

library(flowCore)
library(flowWorkspace)
library(openCyto)
library(ggcyto)
library(flowAI)
library(gridExtra)
library(tidyverse)
library(ggplot2)
```

# Load files

I started with only loading 10 files into the flowset (fs), even though 66 were linked to myfiles

```{r}

setwd("C:/Users/mlb72/OneDrive/Documents/RR")


# Load files #####

myfiles <- list.files(path="C:/Users/mlb72/OneDrive/Documents/RR/RR_R816", # your wd
                      pattern = ".fcs", # file pattern to look for 
                      ignore.case = TRUE) # I have no idea lol

fs <- read.flowSet(myfiles[1:12], # telling how many files at once to load- here, the first 10. These files are HUGE- may help to do in chunks
                   path="C:/Users/mlb72/OneDrive/Documents/RR/RR_R816", # wd 
                   alter.names=TRUE) # no idea 
```

# Extract counts

Here I have to define the parameter we want- B-525-40-A. It's not a main descriptor, so I had to locate the descriptor it is attached to and link them. The final result is being able to use B525-40-A.

I then create a histogram for each file without actually plotting the histogram. I then extract the data into a histogram_list[[i]]. Each file's data is stored in this list.

B-525-40-A is stored as "mid", and counts is stored as "counts"

```{r}
# Extract counts #####

# Inspect parameter descriptions for the first flowFrame
parameter_desc <- pData(parameters(fs[[1]]))$desc
parameter_names <- pData(parameters(fs[[1]]))$name

# Combine descriptions with names for better inspection
param_info <- data.frame(Name = parameter_names, Description = parameter_desc)
print(param_info)

# Assuming you've found the corresponding name for "B-525-40-A" in the desc
desc_of_interest <- "B-525-40-A"

# Find the name corresponding to the description
name_of_interest <- parameter_names[parameter_desc == desc_of_interest]

# Check if the name was found
if (length(name_of_interest) == 0) {
  stop("Description not found in the parameter list.")
} else {
  cat("Mapped parameter description to name:", name_of_interest, "\n")
}

# Initialize a list to store histogram data
histogram_list <- list()

# Loop over each flowFrame in the flowSet
for (i in seq_along(fs)) {
  # Attempt to get the file path for the current flowFrame
  full_file_path <- keyword(fs[[i]], "FILENAME")[[1]]
  
  # Print the full_file_path to verify its content
  print(full_file_path)
  
  # Check if the full_file_path is a character vector
  if (is.character(full_file_path) && length(full_file_path) > 0) {
    # Extract only the file name using basename
    file_name <- basename(full_file_path)
  } else {
    # Handle cases where full_file_path is not valid
    warning(paste("Invalid file path for flowFrame", i, ":", full_file_path))
    file_name <- NA  # Assign NA or some default value
  }
  
  # Print the file name
  cat("Processing file:", file_name, "\n")
  
  # Extract data for the mapped parameter name
  frame_data <- exprs(fs[[i]])[, name_of_interest]
  
  # Generate a histogram without plotting
  hist_data <- hist(frame_data, breaks = 500, plot = FALSE)
  
  # Create a data frame for histogram data
  hist_df <- data.frame(
    mid = hist_data$mids, 
    counts = hist_data$counts
  )
  
  # Add the file name as a single entry (not repeated for each row)
  hist_df$file_name <- file_name[1]
  
  # Store histogram data in the list
  histogram_list[[i]] <- hist_df
}

```

## Check

Check the plot you can build from a file in the histogram_list against an original fs file histogram. It may not be perfect.

```{r}
# Visualize #####

ggplot(histogram_list[[4]], aes(mid, counts))+
  # geom_bar(stat = "identity", width = diff(hist_data$breaks)[1])+
  geom_bar(stat = "identity")+
  labs(title = paste(histogram_list[[1]]$file_name[[1]]))+
  theme_minimal()+
  coord_cartesian(xlim = c(0, 2.5e6))


# check with OG file
ggcyto(fs[[4]], aes(x = `B-525-40-A`))+
  geom_histogram(bins = 500)+
  coord_cartesian(xlim = c(0, 2.5e6))+
  theme_minimal()

```

## Transform

Transform the x, "mid", to log scale

```{r}
# Log scale #####

for (i in seq_along(histogram_list)) {
  histogram_list[[i]]$mid_log <- log(histogram_list[[i]]$mid)
}
```

# Normalize

### To peak height

During the washing steps of the staining protocol, we loose some cells. For this code, we decided to normalize to the max peak to account for this. .

```{r}

# Initialize a list to store normalized histogram data for each flowFrame
normalized_histogram_list <- list()

# Normalize each histogram using the extracted histogram data
for (i in seq_along(histogram_list)) {
  # Get the histogram data frame
  hist_df <- histogram_list[[i]]
  
  # Normalize counts to the highest peak
  hist_df$normalized_counts <- hist_df$counts / max(hist_df$counts)
  
  # Store the normalized histogram data in the list
  normalized_histogram_list[[i]] <- hist_df
}


```

## X axis normalize

```{r}
# Function to adjust x-axis so the peak is at x = 10
adjust_x_axis_for_peak <- function(histogram_data) {
  # Find the x-value (mid_log) where the normalized_counts is max
  peak_x <- histogram_data$mid_log[which.max(histogram_data$normalized_counts)]
  
  # Calculate the shift needed to move the peak to x = 10
  shift <- 10 - peak_x
  
  # Apply the shift to mid_log values
  histogram_data$mid_log_adjusted <- histogram_data$mid_log + shift
  
  return(histogram_data)
}
# Apply the adjustment to each histogram in the list
adjusted_histogram_list <- lapply(normalized_histogram_list, adjust_x_axis_for_peak)

# Optional: Plotting to verify the results
plot_adjusted_histograms <- function(histogram_data_list) {
  # Combine all histograms into one data frame for plotting
  combined_data <- do.call(rbind, lapply(seq_along(histogram_data_list), function(i) {
    df <- histogram_data_list[[i]]
    df$histogram_id <- i
    return(df)
  }))
  
  ggplot(combined_data, aes(x = mid_log_adjusted, y = normalized_counts, color = as.factor(histogram_id))) +
    geom_line() +
    theme_minimal() +
    labs(title = "Adjusted Normalized Histograms",
         x = "Adjusted Log Transformed X-axis",
         y = "Normalized Frequency",
         color = "Histogram ID")
}# Plot the adjusted histograms
plot_adjusted_histograms(adjusted_histogram_list)
```

Our final list is adjusted_histogram_list[[i]] \#################

# Density Curve for nice plots

### Step 1: Cut data

I decided to trim the data to get rid of the outliers to the left of the peaks.

```{r}

min <- 9.5
max <- 11

hist_cut <- list()

for (i in seq_along(adjusted_histogram_list)) {
  df <- adjusted_histogram_list[[i]]
  
  hist_cut[[i]] <- adjusted_histogram_list[[i]] %>%
    filter(mid_log_adjusted >= min, mid_log_adjusted <= max)
}

plots <- function(hist_cut) {
  # Combine all histograms into one data frame for plotting
  combined_data <- do.call(rbind, lapply(seq_along(hist_cut), function(i) {
    df <- hist_cut[[i]]
    df$histogram_id <- i
    return(df)
  }))
  
  ggplot(combined_data, aes(x = mid_log_adjusted, y = normalized_counts, color = as.factor(histogram_id))) +
    geom_line() +
    theme_minimal() +
    labs(title = "Adjusted Normalized Histograms",
         x = "Adjusted Log Transformed X-axis",
         y = "Normalized Frequency",
         color = "Histogram ID")
}# Plot the adjusted histograms
plot_adjusted_histograms(hist_cut)
```

## Step 2: Calculate density

Data stored in 'density_list[[i]]'

```{r}
# Initialize a list to store density data for each flowFrame
density_list <- list()

# Loop through each normalized histogram to calculate density
for (i in seq_along(hist_cut)) {
  # Extract histogram data for the current flowFrame
  hist_df <- hist_cut[[i]]
  filename <- hist_cut[[i]]$file_name[1]
  
  # Compute the density using the midpoints weighted by the normalized counts
  density_data <- density(rep(hist_df$mid_log_adjusted, times = round(hist_df$normalized_counts * 1000)),  # Weighting mids by normalized counts
                          kernel = "gaussian",  # Optional: specify the kernel type
                          bw = "nrd0")  # Optional: bandwidth selection
  
  # Create a data frame for density data
  density_df <- data.frame(
    x = density_data$x,
    y = density_data$y / max(density_data$y),
    file_name = filename# Normalize the density to match histogram normalization
  )
  
  # Add the filename to the density_df
  # Assuming filename is consistent within each adjusted_histogram_list entry
  # density_df$file_name <- unique(hist_df$FILENAME)
  
  # Store the density data in the list
  density_list[[i]] <- density_df
}

```

## Check

```{r}

ggplot(hist_cut[[1]], aes(x = mid_log_adjusted, y = normalized_counts))+
  geom_bar(stat = "identity", fill = "white", alpha = 0)+
  geom_line(data = density_list[[1]], aes(x = x, y = y), col = "black", size =1)+
  theme_classic()+
  coord_cartesian(xlim = c(9, 12))+
  labs(x = "B-525-40-A", y = "Counts", title = density_list[[1]]$file_name)+
  theme(text = element_text(family = "serif"))


```

## Plot all

```{r}
# Initialize a list to store ggplot objects
plot_list <- list()

# Loop through each normalized histogram and create a plot with density curve
for (i in seq_along(hist_cut)) {
  # Extract histogram and density data for the current flowFrame
  hist_df <- hist_cut[[i]]
  density_df <- density_list[[i]]
  file_name <- hist_cut[[i]]$FILENAME
  
  # Create a ggplot object for the current histogram with density curve
  p <- ggplot(hist_df, aes(x = mid_log_adjusted, y = normalized_counts)) +
    geom_bar(stat = "identity", width = diff(hist_df$mid_log_adjusted)[1], fill = "white", alpha = 0) +
    geom_line(data = density_df, aes(x = x, y = y), color = "black", size = 1) +
    labs(x = desc_of_interest, y = "Normalized Frequency") +
    theme_classic() +
    theme(text = element_text(family = "serif")) +
    # theme(panel.border = element_rect(color = "black", fill = NA, linewidth = 0.5))+
    coord_cartesian(xlim = c(9.5, 11))
    
  
  # Add the plot to the list
  plot_list[[i]] <- p
}


# Print all plots
plot_list
```

```{r}
# Initialize a list to store ggplot objects
plot_list <- list()

# Loop through each normalized histogram and create a plot with density curve
for (i in seq_along(hist_cut)) {
  # Extract histogram and density data for the current flowFrame
  hist_df <- hist_cut[[i]]
  density_df <- density_list[[i]]
  file_name <- hist_df$FILENAME[1]  # Extract the file name from the first row
  
  # Create a ggplot object for the current histogram with density curve
  p <- ggplot(hist_df, aes(x = mid_log_adjusted, y = normalized_counts)) +
    geom_bar(stat = "identity", width = diff(hist_df$mid_log_adjusted)[1], fill = "white", alpha = 0) +
    geom_line(data = density_df, aes(x = x, y = y), color = "black", size = 1) +
    labs(x = desc_of_interest, y = "Normalized Frequency", title = file_name) +
    theme_classic() +
    theme(text = element_text(family = "serif")) +
    # theme(panel.border = element_rect(color = "black", fill = NA, linewidth = 0.5))+
    coord_cartesian(xlim = c(9, 12))
    
  # Add the plot to the list
  plot_list[[i]] <- p
}

# Print all plots
plot_list

```

## Export

```{r}

output <- "C:/Users/mlb72/OneDrive/Documents/Rice/Lab/R for students/Rocky_Ren_Summer2024/RR_extra_plots/"
#



for (i in seq_along(plot_list)) {
  plot_name <- paste0("plot_", i, plot_list[[i]]$data$file_name[1], ".png")
  ggsave(filename = file.path(output, plot_name), plot = plot_list[[i]])
}


```

# Fitting

Here I split the peaks by eye- this will need to be adjusted based on the data. I then fit the two, now separate peaks, with a Gaussian function.

## Step 1: Split the curves apart

```{r}

# Step 1: Separate curves #####
## Based on eyeballing general curves

### A: Seperate both curves =====


left_bound <- 9.6
right_bound <- 10.35

left_bound2 <- 10.35
right_bound2 <- 11

## B: Initiate empty lists =====

first_curve_list <-  list()
second_curve_list <- list()

## C: Loop along to split =====

for (i in seq_along(adjusted_histogram_list)) {
  # First curve
  first_curve <- adjusted_histogram_list[[i]] %>%
    filter(mid_log_adjusted > left_bound, mid_log_adjusted < right_bound)
  # Second curve
  second_curve <- adjusted_histogram_list[[i]] %>%
    filter(mid_log_adjusted > left_bound2, mid_log_adjusted < right_bound2)
  
  # Extract the data
  first_curve_list[[i]] <- first_curve
  second_curve_list[[i]] <- second_curve

  }


```

### Plot the curves to check!

```{r}

# First curve #####

half_first_curve_plots <- list()

for (i in seq_along(first_curve_list)) {
  p <- ggplot(first_curve_list[[i]], aes(mid_log_adjusted, normalized_counts))+
    geom_point()
  half_first_curve_plots[[i]] <- p
  
}

print(half_first_curve_plots)

```

```{r}

# Second curve #####

half_second_curve_plots <- list()

for (i in seq_along(second_curve_list)) {
  p <- ggplot(second_curve_list[[i]], aes(mid_log_adjusted, normalized_counts))+
    geom_point()
  half_second_curve_plots[[i]] <- p
  
}

print(half_second_curve_plots)
```

## Step 2: Fit peaks

For both curves, starting parameters need to be defined. Refer to the split curve plots above to estimate a (should be peak max), b (midpoint on the x (always 10 for the first curve), and c (spread).

```{r}
# Step 2: Fit the curves #####

## A: Fit first curve ======

first_fit <- list()

for (i in seq_along(first_curve_list)) {
  x_1 <- first_curve_list[[i]]$mid_log_adjusted
  y_1 <- first_curve_list[[i]]$normalized_counts
  curve <- first_curve_list[[i]]
  
  # Define control parameters for nls
  control <- nls.control(minFactor = 1e-5)
  
  # Try different starting values if necessary
  start_vals <- list(a = max(curve$normalized_counts), b = mean(x_1), c = sd(x_1))
  
  gaussian_model_first <- try(nls(
    y_1 ~ a * exp(-(x_1 - b)^2 / (2 * c^2)),
    start = start_vals,
    control = control
  ), silent = TRUE)
  
  if (inherits(gaussian_model_first, "try-error")) {
    print(paste("Error fitting curve", i))
    next
  }
  
  fit_first_peak <- predict(gaussian_model_first)
  
  fit <- as.data.frame(fit_first_peak)
  
  fit2 <- bind_cols(curve, fit)
  
  first_fit[[i]] <- fit2
}

### Check! -----

ggplot(first_fit[[30]], aes(mid_log_adjusted, normalized_counts))+
  geom_point()+
  geom_line(aes(y = fit_first_peak), linetype = "solid", col = "blue")


## B: Fit second curve =====

second_fit <- list()

for (i in seq_along(second_curve_list)) {
  x_1 <- second_curve_list[[i]]$mid_log_adjusted
  y_1 <- second_curve_list[[i]]$normalized_counts
  curve <- second_curve_list[[i]]
  
  gaussian_model_second <- nls(
    y_1 ~ a * exp(-(x_1 - b)^2 / (2 * c^2)),
    start = list(a = max(curve$normalized_counts), b = 10.62, c = sd(curve$normalized_counts))
  )
  
  fit_second_peak <- predict(gaussian_model_second)
  
  # summary(gaussian_model_second)[1]
  
  fit <- as.data.frame(fit_second_peak)
  
  fit2 <- bind_cols(curve, fit)
  
  second_fit[[i]] <- fit2
}

### Check! -----

ggplot(second_fit[[20]], aes(mid_log_adjusted, normalized_counts))+
  geom_point()+
  geom_line(aes(y = fit_second_peak), linetype = "solid", col = "blue")
```

### Plot all fits to check

```{r}

# First curve #####

first_curve_plots <- list()

for (i in seq_along(first_fit)) {
  p <- ggplot(first_fit[[i]], aes(x = mid_log_adjusted, y = normalized_counts))+
    geom_point()+
    geom_line(aes(y = fit_first_peak), col = "blue")+
    labs(x = desc_of_interest, y = "Counts",
         title = first_fit[[i]]$file_name[1])+
    theme_minimal()
  
  first_curve_plots[[i]] <- p
}

print(first_curve_plots)

```

Second curve plots

```{r}

# Second curve #####

second_curve_plots <- list()

for (i in seq_along(second_fit)) {
  p <- ggplot(second_fit[[i]], aes(x = mid_log_adjusted, y = normalized_counts))+
    geom_point()+
    geom_line(aes(y = fit_second_peak), col = "blue")+
    labs(x = desc_of_interest, y = "Counts",
         title = second_fit[[i]]$file_name[1])+
    theme_minimal()
  
  second_curve_plots[[i]] <- p
}

print(second_curve_plots)

```

### Step 3: Cut curves in half

Here I cut the data in half based on the peak height of the fitted data. We chose to do this because of the potential overlap of the curves in the flow cytometry data. We cut them in half, then mirror the data to create peaks specific to cell cycle phase.

```{r}

# Step 3: Cut curves in half #####

## A: First curve =====

# first_peak_approx_x <- list()

left_of_first <- list()

for (i in seq_along(first_fit)) {
  peak <- max(first_fit[[i]]$fit_first_peak)
  approx_x <- approx(first_fit[[i]]$fit_first_peak, first_fit[[i]]$mid_log_adjusted, xout = peak)$y
  
  # first_peak_approx_x[[i]] <- approx_x
  
  left_of_first[[i]] <- first_fit[[i]] %>%
    filter (mid_log_adjusted <= approx_x)

}

### Check! -----

ggplot(left_of_first[[1]], aes(mid_log_adjusted, normalized_counts))+
  geom_point()+
  geom_line(aes(y = left_of_first[[1]]$fit_first_peak))

## B: Second peak =====

right_of_second <- list()

for (i in seq_along(second_fit)) {
  peak <- max(second_fit[[i]]$fit_second_peak)
  approx_x <- approx(second_fit[[i]]$fit_second_peak, second_fit[[i]]$mid_log_adjusted, xout = peak)$y
  
  # first_peak_approx_x[[i]] <- approx_x
  
  right_of_second[[i]] <- second_fit[[i]] %>%
    filter (mid_log_adjusted >= approx_x)
  
}

### Check! -----

ggplot(right_of_second[[1]], aes(mid_log_adjusted, normalized_counts))+
  geom_point()+
  geom_line(aes(y = right_of_second[[1]]$fit_second_peak))

```

#### Check

Once again plotting all to check!

```{r}

# First curve #####

half_first_plots <- list()

for (i in seq_along(left_of_first)) {
  p <- ggplot(left_of_first[[i]], aes(x = mid_log_adjusted, y = normalized_counts))+
    geom_point()+
    geom_line(aes(y = fit_first_peak), col = "blue")+
    labs(x = desc_of_interest, y = "Counts",
         title = left_of_first[[i]]$file_name[1])+
    theme_minimal()
  
  half_first_plots[[i]] <- p
}

print(half_first_plots)
```

```{r}

# Second curve #####

half_second_plots <- list()

for (i in seq_along(right_of_second)) {
  p <- ggplot(right_of_second[[i]], aes(x = mid_log_adjusted, y = normalized_counts))+
    geom_point()+
    geom_line(aes(y = fit_second_peak), col = "blue")+
    labs(x = desc_of_interest, y = "Counts",
         title = right_of_second[[i]]$file_name[1])+
    theme_minimal()
  
  half_second_plots[[i]] <- p
}

print(half_second_plots)
```

### Step 4: Mirror

As mentioned above, here I mirror the cut data to create a peak exclusive to the cell cycle phase.

```{r}
# Step 4: Mirror fits #####

## A: First curve =====

first_mirror_list <- list()

for (i in seq_along(left_of_first)) {
  mirrored <- left_of_first[[i]]
  mirrored$mid_log_adjusted <- max(left_of_first[[i]]$mid_log_adjusted) + (max(left_of_first[[i]]$mid_log_adjusted) - left_of_first[[i]]$mid_log_adjusted) # Adjust 'mid' for mirroring
  
  first_mirror_list[[i]] <- rbind(left_of_first[[i]], mirrored)
}

### Check! -----

ggplot(first_mirror_list[[10]], aes(mid_log_adjusted, normalized_counts))+
  geom_point()+
  geom_line(aes(y = first_mirror_list[[10]]$fit_first_peak))

## B: Second =====

second_mirror_list <- list()

for (i in seq_along(right_of_second)) {
  mirror_center <- min(right_of_second[[i]]$mid_log_adjusted)
  mirrored2 <- right_of_second[[i]]
  mirrored2$mid_log_adjusted <- mirror_center - (right_of_second[[i]]$mid_log_adjusted - mirror_center)
  
  second_mirror_list[[i]] <- rbind(right_of_second[[i]], mirrored2)
}

### Check! -----

ggplot(second_mirror_list[[20]], aes(mid_log_adjusted, normalized_counts))+
  geom_point()+
  geom_line(aes(y = second_mirror_list[[20]]$fit_second_peak))



```

#### Check

Plot all!

```{r}

# First Curve #####

fill_fitted_first_curve <- list()

for (i in seq_along(first_mirror_list)) {
  p <- ggplot(first_mirror_list[[i]], aes(x = mid_log_adjusted, y = normalized_counts))+
    geom_point()+
    geom_line(aes(y = fit_first_peak), col = "blue")+
    labs(x = desc_of_interest, y = "Counts",
         title = first_mirror_list[[i]]$file_name[1])+
    theme_minimal()
  
  fill_fitted_first_curve[[i]] <- p
}

print(fill_fitted_first_curve)
```

```{r}

# Second Curve #####

fill_fitted_second_curve <- list()

for (i in seq_along(second_mirror_list)) {
  p <- ggplot(second_mirror_list[[i]], aes(x = mid_log_adjusted, y = normalized_counts))+
    geom_point()+
    geom_line(aes(y = fit_second_peak), col = "blue")+
    labs(x = desc_of_interest, y = "Counts",
         title = second_mirror_list[[i]]$file_name[1])+
    theme_minimal()
  
  fill_fitted_second_curve[[i]] <- p
}

print(fill_fitted_second_curve)
```

### Step 5: Integrate area

Now I integrate the area under the fitted peak.

```{r}

# Step 5: Integrate #####

## A: First curve =====

for (i in seq_along(first_mirror_list)) {
  data <- first_mirror_list[[i]]
  
  order1 <- data[order(data$mid_log_adjusted), ]
  
  area_under_curve <- sum(diff(order1$mid_log_adjusted) * (head(order1$fit_first_peak, -1) + tail(order1$fit_first_peak, -1)) / 2)
  
  first_mirror_list[[i]]$area <- area_under_curve
  
}


## B: Second curve =====

for (i in seq_along(second_mirror_list)) {
  data <- second_mirror_list[[i]]
  
  order1 <- data[order(data$mid_log_adjusted), ]
  
  area_under_curve <- sum(diff(order1$mid_log_adjusted) * (head(order1$fit_second_peak, -1) + tail(order1$fit_second_peak, -1)) / 2)
  
  second_mirror_list[[i]]$area <- area_under_curve
  
}



```

### Step 6: Extract Data

Want to extract the data into vectors & bind into new dataframe.

```{r}

# Extract first peak areas #####

first_peak_areas <- rep(0, length(first_mirror_list))

for(i in 1:length(first_mirror_list)) {
  first_peak_areas[[i]] <- first_mirror_list[[i]]$area[1]
}

# Extract second peak areas #####

second_peak_areas <- rep(0, length(second_mirror_list))

for(i in 1:length(second_mirror_list)) {
  second_peak_areas[[i]] <- second_mirror_list[[i]]$area[1]
}

# Extract file names #####

peak_names <- rep("a", length(first_mirror_list))

for(i in 1:length(first_mirror_list)) {
  peak_names[[i]] <- first_mirror_list[[i]]$file_name[1]
}

# Bind data #####

final.file <- as.data.frame(bind_cols(peak_names, first_peak_areas, second_peak_areas))

# View(final.file)

# Update column names #####

final.file$file_name <- final.file$...1
final.file$first_peak_area <- final.file$...2
final.file$second_peak_area <- final.file$...3
```

# Create Stacked Barplot

The final step of this analysis is a stacked barplot. This experiment was done over a 12hr period. File names do not correspond with timing, so we load in a "Match.sheet.csv" with corresponding metadata and match it based on file_name to the extracted data.

**This is where the code breaks down and needs to be organized.**

## Step 1: Load match.sheet.csv

The match sheet contained data regarding time of fixation and replicate.

Technically I did export the data and copy and pasted the metadata in and reloaded. A matching code will be uploaded eventually.

```{r}
# Load match #####

## Set wd =====

# wd(filepath)

match <- read.csv(file = "match.sheet.csv", header = T, sep = ",")

## Organize

# match$Time <- as.POSIXct(match$Time)


```

Side quest: Calc S Phase. S phase lives bewteen the peaks of G1 and G2/ Mitosis. We tried to caluclate it by integrating the area underneath the density curve and subtracting the integrated areas of the two fitted peaks.

```{r}
# Calculate S Phase

# Denisty data =====

density_area_list <- list()

for (i in seq_along(density_list)) {
  data <- density_list[[i]]
  
  order <- data[order(data$x), ]
  
  density_area <- sum(diff(order$x) *(head(order$y, -1) + tail(order$y, -1)) / 2)
  
  density_list[[i]]$area <- density_area
}


# Extract denisty areas #####

density_areas <- rep(0, length(density_list))

for (i in 1:length(density_list)) {
  density_areas[[i]] <- density_list[[i]]$area[1]
}
# bind file =====

final.file <- as.data.frame(bind_cols(peak_names, first_peak_areas, second_peak_areas, density_areas))

# View(final.file)

final.file$file_name <- final.file$...1
final.file$first_peak_area <- final.file$...2
final.file$second_peak_area <- final.file$...3
final.file$density_area <- final.file$...4



final.file$S_phase <- final.file$density_area - final.file$first_peak_area - final.file$second_peak_area

fin <- final.file

write.csv(fin, "C:/Users/mlb72/OneDrive/Documents/RR/RR_168.csv")

# Data #####

RR_R816 <- read.csv(file = "RR_R816.csv", header = T, sep = ",")

RR_816 <- read.csv(file = "RR_618_data.csv", header = T, sep = ",")

RR_R168 <- read.csv(file = "RR_R168.csv", header = T, sep = ",")

RR_cons <- read.csv(file = "RR_cons.csv", header = T, sep = ",")

RR_168 <- read.csv(file = "RR_168.csv", header = T, sep = ",")

```

## Step 2: Average the data

For our experiment structure, we decided to average the replicates.

```{r}

fin <- final.file %>%
  group_by(Replicate) %>%
  summarize(
    across(where(is.numeric),   
           list(mean = ~ mean(., na.rm = TRUE),
                sd = ~ sd(., na.rm = TRUE)),
           .names = "{.col}_{.fn}"),
    across(where(is.character), first),
    Time_mean = as.POSIXct(mean(as.numeric(Time), na.rm = TRUE), origin = "1970-01-01"),
    Time_sd = sd(as.numeric(Time), na.rm = TRUE),
    .groups = 'drop'
  ) %>%
  mutate(Time_sd = as.difftime(Time_sd, units = "secs"))

```

## Step 3: Change the dataframe structure

```{r}
# Step 2: adjust df

adj_file <- fin %>%
  select(file_name, first_peak_area, second_peak_area, S_phase) %>%
  pivot_longer(cols = c(first_peak_area, second_peak_area, S_phase),
               names_to = "area",
               values_to = "value")


```

## Step 4: Plot the stacked barplot

```{r}
# Step 3: Plot!! 


Final_Plot <- ggplot(adj_file, aes(x = Time.After.the.onset.of.light_mean, y = value, fill = area))+
  geom_bar(position = "fill", stat = "identity")+
  # scale_fill_manual(values = c("purple2", "green3", "pink"))+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 0, vjust = 0.5, hjust = 0.5))+
  theme(text = element_text(family = "serif"))+
  labs(y = "Frequency",
       x = "Hours After Light On",
       fill = "Phase:")+
  scale_x_continuous(breaks = c(-2, 0, 2, 4, 6, 8, 10))+
  scale_fill_manual(values = c("darkolivegreen3", "darkseagreen", "forestgreen"),
                    labels = c("G1_mean" = "G1", "G2.Mitosis_mean" = "G2/Mitosis", "S_mean" = "S"))+
  theme(legend.position = "bottom")+
  theme(
    # axis.text.x=element_blank(), #remove x axis labels
    #     axis.ticks.x=element_blank(),
    #     axis.title.x = element_blank(),#remove x axis ticks
        axis.title.x = element_text(size = 14),
        axis.text.x = element_text(size = 12),
        axis.title.y = element_text(size = 14),
        axis.text.y = element_text(size = 12))

print(Final_Plot2)
```

### Final Export

```{r}

ggsave(Final_Plot, 
       filename = "Reversed_816_RR.jpeg",
       device = "jpeg",
       height = 4, width = 8, units = "in")

```

```{r}

write.csv(fin, "C:/Users/mlb72/OneDrive/Documents/RR/RR_618_data.csv")


```
